---
title: "Data Package Checks Report"
rmd_author: "Bibi Powers-McCormack"
date_created: "2024-06-12"
date_updated: "2024-06-12"
output: 
  html_document:
    toc: TRUE
    theme: simplex
---

This report was created on `r Sys.Date()` by `r report_author`.

```{r, echo=FALSE}

# see if the data package passed all checks

# get total number of checks
all_checks_summary <- output$data_checks_summary %>% 
  select(requirement, assessment, source) %>% 
  distinct()

# see if any failed
failed_checks <- output$data_checks_summary %>% 
  select(requirement, assessment, source, pass_check) %>% 
  filter(pass_check == FALSE)

y_required_total <- all_checks %>% 
  filter(requirement == "required") %>% 
  count() %>% 
  pull(n)

x_required_failed <- failed_checks %>% 
  filter(requirement == "required") %>% 
  count() %>% 
  pull(n)

y_recommended_total <- all_checks %>% 
  filter(requirement == "recommended*") %>% 
  count() %>% 
  pull(n)

x_recommended_failed <- failed_checks %>% 
  filter(requirement == "recommended*") %>% 
  count() %>% 
  pull(n)

```

## Summary
**`r basename(output$input$inputs$directory)`  **
  
> <p style='color:black;'><strong>**Your data package passed `r y_required_total - x_required_failed` of `r y_required_total` required test(s).**  </strong></p>
> <p style='color:black;'><strong>**Your data package passed `r y_recommended_total - x_recommended_failed` of `r y_recommended_total` recommended* test(s).** </strong></p>


```{r echo=FALSE}
# display table of checks
all_check_table <- all_checks %>% 
  left_join(failed_checks, by = join_by(requirement, assessment, source)) %>% 
  mutate(pass_check = case_when((pass_check != FALSE | is.na(pass_check)) ~ TRUE, T ~ pass_check))

all_check_table %>% 
  mutate(pass_check = case_when(pass_check == FALSE ~ "FAILED", 
                                pass_check == TRUE ~ "PASSED",
                                T ~ NA_character_)) %>% 
  kable() %>% 
  kable_styling() %>%
  row_spec(which(all_check_table$pass_check == TRUE), color = "gray") %>% 
  row_spec(which(all_check_table$pass_check == FALSE), color = "black", bold = T)
```

## Required Checks {.tabset}
  
### Required Checks
  
The following checks are required to pass before a data package will be published.  
1. **`includes required files`**: Certain files are required and must be named correctly (e.g., readme, flmd, dd).  

### `includes required files` {.tabset}

```{r echo=FALSE}

# determine if this check failed
failed_includes_required_files <- output$data_checks %>% 
  filter(assessment == "includes required files" & pass_check == FALSE) %>% 
  count() %>% 
  pull()

```
> <p style='color:black;'><strong> **Your data package is missing `r failed_includes_required_files` required file(s).** </strong></p>


In addition to your files, we require a file-level metadata file (FLMD), data dictionary (DD), and readme in each data package. To improve interoperability with ESS-DIVE's tools, these file names have specific file naming conventions.  

- FLMD: file name must end with `*flmd.csv`. You may prepend anything to the beginning.  
- DD: file name must end with `*dd.csv`. You may prepend anything to the beginning.  
- readme: file name must begin with `readme*`. You may append anything to the end and the file extension must be non-proprietary.  

```{r echo=FALSE}

# show pie chart of all passed/failed files
output$data_checks %>% 
  filter(assessment == "includes required files") %>% 
  select(pass_check, input) %>% 
  mutate(pass_check = case_when(pass_check == TRUE ~ "PASSED",
                                pass_check == FALSE ~ "FAILED")) %>%  
  count(pass_check) %>% 
  ggplot(aes(x = "", y = n, fill = pass_check)) + 
    geom_bar(sta = "identity", width = 1) + # bar chart with no x axis
    coord_polar(theta = "y") +
    scale_fill_manual(values = c("FAILED" = "red", "PASSED" = "grey")) +
    labs(title = "Evaluation of Required Files") +
    theme_void()

if (failed_includes_required_files != 0) {

  # if there are failed checks, display table
  output$data_checks %>% 
    filter(assessment == "includes required files") %>% 
    select(requirement, pass_check, assessment, value) %>% 
      mutate(pass_check = case_when(pass_check == FALSE ~ "FAILED", 
                                  pass_check == TRUE ~ "PASSED",
                                  T ~ NA_character_)) %>% 
    datatable(rownames = F,
            filter = "top",
            class = "stripe")
  
}

```

## Recommended Checks {.tabset}
  
### Recommended Checks
The following checks are recommended to pass before a data package will be published. Passing `recommended*` checks prepare your data package to be included into ESS-DIVE's Fusion Database, allowing your data and scripts to be searchable within the deep dive API search tool. The RC-SFA project expects that all Study Data Packages adhere to both `required` and `recommended*` checks. Manuscript Data Packages must adhere to the `required` checks and are strongly encouraged to consider the `recommended*` checks.   

Recommended checks:  
1. **`no proprietary files`**: To increase increase accessibility, proprietary files should be converted to non-proprietary options (e.g., .docx are converted to .pdf).  
2. **`no special characters` in directory names**: The only characters allowed in folder names are lowercase letters, uppercase letters, numbers, underscores, periods, and hyphens.  
3. **`no special characters` in file names**: The only characters allowed in file names are lowercase letters, uppercase letters, numbers, underscores, periods, and hyphens.  
4. **`no special characters` in column names**: The only characters allowed in tabular column headers are lowercase letters, uppercase letters, numbers, underscores, periods, and hyphens. It is recommended that columns begin with a letter.  
5. **`no duplicate names` in column names**: Column names should be unique within the same file. The same column name can repeat across files if it shares the same definition and unit.  

### `no proprietary files` {.tabset}

```{r echo=FALSE}

# determine if this check failed
failed_no_proprietary_files <- output$data_checks %>% 
  filter(assessment == "no proprietary files") %>% 
  filter(pass_check == FALSE) %>% 
  count() %>% 
  pull()
```
> <p style='color:black;'><strong> **Your data package has `r failed_no_proprietary_files` proprietary file(s).**  </strong></p>

``` {r echo=FALSE}

# show bar chart of file extensions
output$data_checks %>% 
  filter(assessment == "no proprietary files") %>% 
  select(pass_check, input) %>% 
  mutate(extension = tools::file_ext(input)) %>% 
  mutate(pass_check = case_when(pass_check == TRUE ~ "PASSED",
                                pass_check == FALSE ~ "FAILED")) %>% 
  group_by(extension, pass_check) %>% 
  summarise(extension_count = n(), .groups = "drop") %>% 
  
  ggplot(., aes(x = extension, y = extension_count, fill = pass_check, text = paste("files:", extension_count))) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = extension_count), vjust = -0.5) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    scale_fill_manual(values = c("FAILED" = "red", "PASSED" = "grey")) +
    theme_minimal() +
    labs(x = "File Extension", y = "Count of Files", title = paste0("Evaluation of File Extensions"))
    
if (failed_no_proprietary_files != 0) {
  
  # if there are proprietary extensions, show table of them
output$data_checks %>% 
    select(requirement, pass_check, assessment, value, file) %>%
    mutate(pass_check = as.character(case_when(pass_check == FALSE ~ "FAILED",
                                pass_check == TRUE ~ "PASSED",
                                T ~ NA_character_))) %>%
    filter(assessment == "no proprietary files") %>%
    datatable(rownames = F,
              filter = "top", 
              class = "stripe")
  
}

  
```


### `no special characters` {.tabset}

```{r include=FALSE}

# determine if these checks failed
output$data_checks %>% 
  filter(assessment == "no special characters") %>% 
  filter(pass_check == FALSE) %>% 
  count(source)


```


> <p style='color:black;'><strong> **Your data package has `r failed_no_special_chrs_dirs` folder name(s) that contain special characters.**  </strong></p>
> <p style='color:black;'><strong> **Your data package has `r failed_no_special_chrs_files` file name(s) that contain special characters.**  </strong></p>
> <p style='color:black;'><strong> **Your data package has `r failed_no_special_chrs_cols` file(s) that contain column names with special characters.**  </strong></p>

Allowed characters include lowercase letters, uppercase letters, numbers, underscores, periods, and hyphens. Any other character (e.g., spaces, symbols, etc.) will cause the check to fail. 

#### Folder Names {.tabset}
```{r echo=FALSE}

# if special chrs are present, show figures and tables

# show pie chart of the number of folder names that have a special chr
output$data_checks %>% 
  filter(assessment == "no special characters") %>% 
  filter(source == "directory_name") %>%
  select(pass_check, input) %>% 
  mutate(pass_check = case_when(pass_check == TRUE ~ "PASSED",
                                pass_check == FALSE ~ "FAILED")) %>%  
  count(pass_check) %>% 
  ggplot(aes(x = "", y = n, fill = pass_check)) + 
    geom_bar(sta = "identity", width = 1) + # bar chart with no x axis
    coord_polar(theta = "y") +
    scale_fill_manual(values = c("FAILED" = "red", "PASSED" = "grey")) +
    labs(title = "Evaluation of Folders with Special Characters") +
    theme_void()

# show a bar chart of the special chrs present in the data package
output$data_checks %>% 
  filter(assessment == "no special characters") %>% 
  filter(source == "directory_name") %>%
  select(pass_check, input, value) %>% 
  mutate(pass_check = case_when(pass_check == TRUE ~ "PASSED",
                                pass_check == FALSE ~ "FAILED")) %>% 
  group_by(value, pass_check) %>% 
  summarise(special_chr_count = n(), .groups = "drop") %>% 
  
  ggplot(., aes(x = value, y = special_chr_count, fill = pass_check, text = paste("folders:", special_chr_count))) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = special_chr_count), vjust = -0.5) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    scale_fill_manual(values = c("FAILED" = "red", "PASSED" = "grey")) +
    theme_minimal() +
    labs(x = "Special Characters", y = "Count of Folders", title = paste0("Evaluation of Special Characters in Folders"))


# show table



```


#### File Names {.tabset}

#### Column Names {.tabset}

### `no duplicate names` {.tabset}

## About Your Data {.tabset}

### Directory Tree

Below is a list of all of your folders and files that were present when the report was generated.
```{r, echo = FALSE}
dir_tree(output$input$inputs$directory, recurse = T)
```

### Tabular Data {.tabset}

## Full Data Checks Report {.tabset}

### Summary Table {.tabset}
```{r echo=FALSE}
output$data_checks_summary %>% 
    mutate(pass_check = case_when(pass_check == TRUE ~ "PASSED",
                                pass_check == FALSE ~ "FAILED")) %>% 
      datatable(rownames = F,
              filter = "top", 
              class = "stripe")
```


### Data Checks {.tabset}
```{r echo=FALSE}
output$data_checks %>% 
    mutate(pass_check = case_when(pass_check == TRUE ~ "PASSED",
                                pass_check == FALSE ~ "FAILED")) %>% 
      datatable(rownames = F,
              filter = "top", 
              class = "stripe")
```
















