---
title: "Data Package Checks Report"
rmd_author: "Bibi Powers-McCormack"
date_created: "2024-06-12"
date_updated: "2025-02-04"
output: 
  html_document:
    toc: FALSE
    theme: simplex
---

This report was created on `r Sys.Date()` by `r report_author`.

```{r, echo=FALSE}
# see if the data package passed all checks

# script inputs: 
  # requires `output` list from the check_data_package() function

# get total number of checks
all_checks_summary <- data_package_checks$data_checks_summary %>% 
  select(requirement, assessment, source) %>% 
  distinct()

# see if any failed
failed_checks <- data_package_checks$data_checks_summary %>% 
  select(requirement, assessment, source, pass_check) %>% 
  filter(pass_check == FALSE)

y_required_total <- all_checks_summary %>% 
  filter(requirement == "required") %>% 
  count() %>% 
  pull(n)

x_required_failed <- failed_checks %>% 
  filter(requirement == "required") %>% 
  count() %>% 
  pull(n)

y_recommended_total <- all_checks_summary %>% 
  filter(requirement == "strongly_recommended") %>% 
  count() %>% 
  pull(n)

x_recommended_failed <- failed_checks %>% 
  filter(requirement == "strongly_recommended") %>% 
  count() %>% 
  pull(n)

```

<p style='color: black;'> Data Package: **`r basename(data_package_checks$input$inputs$directory)`** </p>
  
> <p style='color: `r ifelse(x_required_failed > 0, "lightcoral", "green")`;'><strong>**Your data package passed `r y_required_total - x_required_failed` of `r y_required_total` required test(s).**  </strong></p>
> <p style='color: `r ifelse(x_recommended_failed > 0, "lightcoral", "green")`;'><strong>**Your data package passed `r y_recommended_total - x_recommended_failed` of `r y_recommended_total` strongly_recommended test(s).** </strong></p>


## Did your data package pass? {.tabset}  
**Each tab provides an evaluation of specific data quality checks designed to ensure compliance with best practices for accessibility and usability. Review any flagged items in the visualizations or tables, make corrections as needed (e.g., add missing files, replace proprietary files, rename columns, or remove duplicates), and re-run the checks to confirm compliance.**   

### Overview  
- Required checks are required to pass before a data package will be published.  
- Recommended checks are suggested to pass before a data package will be published. Passing `strongly_recommended` checks prepares your data package to be included into ESS-DIVE's Fusion Database, allowing your data and scripts to be searchable within the deep dive API search tool.  
- The RC-SFA project expects that all Study Data Packages adhere to both `required` and `strongly_recommended` checks. Manuscript Data Packages must adhere to the `required` checks and are encouraged to consider the `strongly_recommended` checks.  

```{r echo=FALSE}
# display table of checks

data_package_checks$data_checks_summary %>% 
  select(pass_check, assessment, requirement, source, file_count) %>% 

  # prep up pass_check colors
  mutate(pass_check = case_when(pass_check == FALSE ~ "FAILED", 
                              pass_check == TRUE ~ "PASSED",
                              T ~ NA_character_)) %>%
  mutate(pass_check = cell_spec(pass_check, background = case_when(pass_check == "PASSED" ~ "lightgreen",
                                                                  pass_check == "FAILED" ~ "lightcoral"))) %>% 

  # make nice table
  kable("html", escape = FALSE) %>% 
  kable_styling("striped", full_width = FALSE, position = "center") %>% 
  row_spec(which(data_package_checks$data_checks_summary$pass_check == TRUE), color = "gray") %>% 
  row_spec(which(data_package_checks$data_checks_summary$pass_check == FALSE), color = "black", bold = T)


```

---

### `includes required files` {.tabset}

```{r echo=FALSE}

# determine if this check failed
failed_includes_required_files <- data_package_checks$data_checks %>% 
  filter(assessment == "includes required files" & pass_check == FALSE) %>% 
  count() %>% 
  pull()

```
> <p style='color: `r ifelse(failed_includes_required_files > 0, "lightcoral", "green")`;'><strong> **Your data package is missing `r failed_includes_required_files` required file(s).** </strong></p>


In addition to your files, we require a file-level metadata file (FLMD), data dictionary (DD), and readme in each data package Proprietary files should be converted to non-proprietary options (e.g., .docx are converted to .pdf). To improve interoperability with ESS-DIVE's tools, these file names have specific file naming conventions.  

- **FLMD**: file name must end with `*flmd.csv`. You may prepend anything to the beginning.  
- **DD**: file name must end with `*dd.csv`. You may prepend anything to the beginning.  
- **readme**: file name must begin with `readme*` and end with `.pdf`. You may append anything between "readme" and ".pdf".  

```{r echo=FALSE}

# show pie chart of all passed/failed files
data_package_checks$data_checks %>% 
  filter(assessment == "includes required files") %>% 
  select(pass_check, value, input) %>% 
  mutate(pass_check = case_when(pass_check == TRUE ~ "PASSED",
                                pass_check == FALSE ~ "FAILED")) %>%  
  mutate(value = case_when(value == ".*flmd\\.csv$" ~ "FLMD", 
                           value == ".*dd\\.csv$"  ~ "DD",
                           value == "^readme.*\\.pdf$" ~ "readme")) %>% 
  mutate(pass_check = cell_spec(pass_check, background = case_when(pass_check == "PASSED" ~ "lightgreen",
                                                                   pass_check == "FAILED" ~ "lightcoral"))) %>% 
  rename(required_file = value) %>% 
  kable("html", escape = FALSE) %>% 
  kable_styling("striped", full_width = FALSE)

```

---


### `no proprietary files` {.tabset}

```{r echo=FALSE}

# determine if this check failed
failed_no_proprietary_files <- data_package_checks$data_checks %>% 
  filter(assessment == "no proprietary files") %>% 
  filter(pass_check == FALSE) %>% 
  count() %>% 
  pull()
```
> <p style='color: `r ifelse(failed_no_proprietary_files > 0, "lightcoral", "green")`;'><strong> **Your data package has `r failed_no_proprietary_files` proprietary file(s).**  </strong></p>

ESS-DIVE requests that all files are saved in a non-proprietary file format. While not exhaustive, this check confirms that the following common proprietary extensions are not included: `r input_parameters$non_proprietary_extensions`

``` {r echo=FALSE}

# show bar chart of file extensions
data_package_checks$data_checks %>% 
  filter(assessment == "no proprietary files") %>% 
  select(pass_check, input) %>% 
  mutate(extension = tools::file_ext(input)) %>% 
  mutate(pass_check = case_when(pass_check == TRUE ~ "PASSED",
                                pass_check == FALSE ~ "FAILED")) %>% 
  group_by(extension, pass_check) %>% 
  summarise(extension_count = n(), .groups = "drop") %>% 
  
  ggplot(., aes(x = extension, y = extension_count, fill = pass_check, text = paste("files:", extension_count))) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = extension_count), vjust = -0.5) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
    scale_fill_manual(values = c("FAILED" = "darkred", "PASSED" = "grey")) +
    theme_minimal() +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    theme(panel.grid.minor = element_blank()) +
    labs(x = "File Extension", 
         y = "Count of Files", 
         title = "Evaluation of File Extensions",
         subtitle = "Red indicates the presence of proprietary files")

  
```

---

### `no special characters` {.tabset}

```{r include=FALSE}

# determine if these checks failed
failed_no_special_chrs <- data_package_checks$data_checks %>%
  filter(assessment == "no special characters") %>% 
  filter(pass_check == "FALSE") %>% 
  count(source, pass_check) %>% 
  summarise(n = sum(n)) %>% 
  pull()

```


> <p style='color: `r ifelse(failed_no_special_chrs > 0, "lightcoral", "green")`;'><strong> **Your data package has `r failed_no_special_chrs` folder/file/column name(s) that contain special characters.**  </strong></p>  

Special characters are not allowed in folder, file, or column names and will cause the check to fail.  

- **`no special characters` in directory names**: The only characters allowed in folder names are lowercase letters, uppercase letters, numbers, underscores, periods, and hyphens.  
- **`no special characters` in file names**: The only characters allowed in file names are lowercase letters, uppercase letters, numbers, underscores, periods, and hyphens.  
- **`no special characters` in column names**: The only characters allowed in tabular column headers are lowercase letters, uppercase letters, numbers, underscores, periods, and hyphens. It is recommended that columns begin with a letter.  

```{r echo=FALSE}

# show bar chart of the count of files/folders/cols that have special chrs
data_package_checks$data_checks %>% 
  filter(assessment == "no special characters") %>% 
  select(pass_check, source, file) %>%
  distinct() %>% 
  group_by(source, file) %>% 
  summarise(pass_check = !any(!pass_check), .groups = "drop") %>% # collapsing so if a file has some cols that pass and others that don't it counts the file as having failed the check
  group_by(pass_check, source) %>% 
  summarise(file_count = n(), .groups = "drop") %>% 
  mutate(pass_check = case_when(pass_check == TRUE ~ "PASSED",
                                pass_check == FALSE ~ "FAILED")) %>%
  
  
  ggplot(., aes(x = source, y = file_count, fill = pass_check, text = paste("files:", file_count))) +
    geom_bar(stat = "identity") +
    geom_text(aes(label = file_count), position = position_stack(vjust = 0.5)) +
    scale_fill_manual(values = c("FAILED" = "darkred", "PASSED" = "grey")) +
    theme_minimal() +
    scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +
    theme(panel.grid.minor = element_blank()) +
    labs(x = "Source", 
         y = "Count of Files/Folders", 
         title = "Evaluation of Files and Folders with Special Characters",
         subtitle = "Red indicates the presence of special characters")


```

---

### `no duplicate names` {.tabset}

```{r include=FALSE}

# determine if these checks failed
failed_duplicate_names <- data_package_checks$data_checks %>%
  filter(assessment == "no duplicate names") %>% 
  filter(pass_check == FALSE) %>% 
  select(file) %>% 
  distinct() %>% 
  count() %>% 
  pull(n)

```
> <p style='color: `r ifelse(failed_duplicate_names > 0, "lightcoral", "green")`;'><strong> **Your data package has `r failed_duplicate_names` file(s) that contain duplicate column names.**  </strong></p>  

Column names should be unique within each file. Column names can be repeated across files if they share the same definition.

```{r echo=FALSE}

data_package_checks$tabular_report %>% 
  group_by(file_name, column_name) %>%
  summarise(column_counts_within_files = as.integer(n()), .groups = "drop") %>%
  group_by(column_name) %>% 
  mutate(column_counts_between_files = as.integer(n())) %>% 
  mutate(pass_check = case_when(column_counts_within_files > 1 ~ "FAILED", 
                                column_counts_between_files > 1 ~ "WARNING", 
                                T ~ "PASSED")) %>% 
  ungroup() %>% 
  mutate(column_name = factor(column_name, levels = rev(sort(unique(column_name))))) %>% 

  ggplot(aes(x = column_name, y = column_counts_within_files, fill = pass_check)) +
    geom_bar(stat = "identity", position = "stack") +
    scale_fill_manual(values = c("FAILED" = "darkred", "WARNING" = "#e3a84f","PASSED" = "grey")) +
    scale_y_continuous(breaks = scales::pretty_breaks()) +
    labs(x = "Column Name", 
         y = "Count", 
         title = "Column names across all files",
         subtitle = paste0("Red (failed) = column name appears more than once within the same file.", "\n", "Orange (warning) = column name appears more than once across files.")) +
    theme_minimal() +
    theme(panel.grid.minor = element_blank()) +
    coord_flip()

```

---

### Tabular Data {.tabset}

```{r include = FALSE}

# determine if there are missing or negative values

data_are_missing <- data_package_checks$tabular_report %>% 
  filter(num_missing_rows > 0) %>% 
  count() %>% 
  pull()

data_are_negative <- data_package_checks$tabular_report %>% 
  filter(num_negative_rows > 0) %>% 
  count() %>% 
  pull()

```

> <p style='color: `r ifelse(data_are_missing > 0, "lightcoral", "green")`;'><strong> **Number of missing values in your data package: `r data_are_missing`**  </strong></p>  
> <p style='color: `r ifelse(data_are_negative > 0, "lightcoral", "green")`;'><strong> **Number of negative values in your data package: `r data_are_negative`**  </strong></p>  

This table summarizes column-level details across all the tabular files in your data package, highlighting potential issues in your data.    

Here are a few key columns to check:  
- `num_missing_values`: check for missing data (the value should be 0 unless missing values are expected).  
- `range_min` & `range_max`: if applicable, validate numeric ranges to ensure realistic values.  
- `num_negative_values`: if applicable, check for unexpected negative values.  
- `data_preview`: verify the top values in each column; this displays the top 5 most frequent values for each column.  


```{r echo=FALSE}

data_package_checks$tabular_report %>% 
  select(file_name, column_name, column_type, num_rows, num_unique_rows, num_missing_rows, num_negative_rows, range_min, range_max, top_counts) %>% # reorder
  rename(num_missing_values = num_missing_rows,
         num_negative_values = num_negative_rows,
         data_preview = top_counts) %>% 
  # convert dbl to int and chr to fct
   mutate(across(
    .cols = everything(),
    .fns = ~ {
      if (is.double(.)) {
        as.integer(.)
      } else if (is.character(.)) {
        as.factor(.)
      } else {
        .
      }
    }
  )) %>% 
    datatable(options = list(autoWidth = TRUE,
                     scrollX = TRUE,
                     columnDefs = list(list(width = '100%', targets = "_all"),
                                       # hover text for top_counts
                                       list(targets = 9,  # Index of the column (0-based) where you don't want text wrapping
                                            render = JS("function(data, type, row, meta) { return '<div style=\"white-space: nowrap; overflow: hidden; text-overflow: ellipsis;\" title=\"' + data + '\">' + data + '</div>';
           }"
        )))),
      rownames = F,
      filter = "top",
      class = "stripe row-border") %>% 
  
  # color missing values red
      formatStyle(c("num_missing_values", "num_negative_values"), # Column to apply the style
                  color = styleInterval(0, c(NA, 'red'))  # Highlight cells with value > 0 in red
                  )

```

---


## Detailed Error Report {.tabset}

The Summary Table and Data Checks tabs provide high-level and detailed insights into the quality of your data. Use the summary to identify major issues and the detailed checks to troubleshoot specific problems. 

1. Start with Summary Table. Focus on checks with `pass_check = FAILED` to locate critical issues first.  
2. Then use the Data Checks tab to find problematic files and review detailed explanations of failures.  
3. Correct the issues in your data files based on the failures identified, then re-run the checks to ensure all problems have been resolved.   

### Failed Checks {.tabset}
```{r echo=FALSE}
data_package_checks$data_checks %>% 
    filter(pass_check == FALSE) %>% 
    select(assessment, requirement, input, source, value, file) %>% 
    mutate(across(c(assessment, requirement, source, file), as.factor)) %>% 
      datatable(rownames = F,
              filter = "top", 
              class = "stripe")
```

---

### Passed Checks {.tabset}
```{r echo=FALSE}
 data_package_checks$data_checks %>% 
    filter(pass_check == TRUE) %>% 
    select(assessment, requirement, input, source, value, file) %>% 
    mutate(across(c(assessment, requirement, source, file), as.factor)) %>% 
      datatable(rownames = F,
              filter = "top", 
              class = "stripe")

```

---

### Directory Tree {.tabset}

Below is a list of all of your folders and files that were present when the report was generated.
```{r, echo = FALSE}
dir_tree(data_package_checks$input$inputs$directory, recurse = T)
```

---
